---
title: "final Project"
author: "Elisabeth Balke, Caleb Purcell, Kordell Schrock, Patrick Wenzel"
date: "11/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(dplyr.summarise.inform = FALSE)
options(dplyr.tibble.inform = FALSE)
library (rtweet)
library(twitteR)
library(ggplot2)
library(stringr)
library(tidyr)
library(dplyr)
library(tidytext)
library(tidyverse)
library(textdata)
#The get_sentiments function returns a tibble, so to take a look at what is included as “positive” and “negative” sentiment, you will need to filter. from 
library(widyr)
library(igraph)
library(ggraph)
library(data.table)
library(lubridate)
library(readxl)
library(expss)
library(ggrepel)
library(plotly)
poll <- read.csv("./Data/statepredictions.csv")
colnames(poll) <- c('state', 'trumpPercent', 'bidenPercent', 'otherPercent', 'electoralVotes')
```

```{r, message=FALSE, warning=FALSE}
consumer_key <- "Mt2dzgMAOvWqU13zmdtAxovNR"
consumer_secret <- "1CYFDynCiKvmUikEfZYffHuX1QJwqxNwko4liZmQxVXtj9N7Gk"
access_token <- "1095430370198208512-1IyFwJEHPJpFVbGinTarjXapjQjxV3"
access_token_secret <- "5McWrmhfBedpJMHZSDqoTMk9sGYLqrxv4MuZB0w132mIp"

#Create Twitter Connection
setup_twitter_oauth(consumer_key, consumer_secret, access_token, access_token_secret)

account <- "realdonaldtrump"
account.timeline <- userTimeline(account, n=10, includeRts = TRUE)
```

```{r, message=FALSE, warning=FALSE}
print(account.timeline)
```

```{r, message=FALSE, warning=FALSE}
app_name ="Kordell's app to connect to R"
create_token(app = app_name,
             consumer_key = consumer_key,
             consumer_secret = consumer_secret,
             access_token = access_token,
             access_secret = access_token_secret)
```

```{r eval = FALSE, echo = FALSE, message=FALSE, warning=FALSE}

trumpHashtags <- search_tweets(
"#DonaldTrump", n = 10000 , retryonratelimit = TRUE
)
fwrite(trumpHashtags, file ="./Data/donaldTrump_TwitterAPI.csv")
```

```{r eval = FALSE, echo = FALSE, message=FALSE, warning=FALSE}
bidenHashtags <- search_tweets(
"#JoeBiden", n = 10000 , retryonratelimit = TRUE
)
fwrite(bidenHashtags, file ="./Data/joeBiden_TwitterAPI.csv")
```

```{r, message=FALSE, warning=FALSE}
trumpHashtags <- readr::read_csv("./Data/donaldTrump_TwitterAPI.csv")
search_term <- "#DonaldTrump"
trumpHashtags$date <- substr(trumpHashtags$created_at,1,10)
trumpHashtags <- trumpHashtags[trumpHashtags$date == '2020-11-16',]
print(trumpHashtags)
```

### 1.What is the current sentiment for Biden and Trump from regular twitter users? Are the tweets positive, netural or negative in sentiment? 

    We found that from getting twitter data instantly from the Twitter API on Nov 16 that Biden was more postive in tweets. While Trump was more negative in tweets. This could be from Biden winning the elction and Trump losing the election. 

```{r, message=FALSE, warning=FALSE}
by <- 'hour'
rtweet::ts_plot(trumpHashtags, by = by, trim = 1, col = "red") + geom_point() + 
theme_minimal() + labs(title = paste0('Tweets mentioning "',
search_term,'" by ',by),
x = 'Date', y = 'Count', caption = 'Source: Twitter API 10k Tweets')
ggsave("./Pictures/trumpByHour.png")
```

```{r, message=FALSE, warning=FALSE}
sentiment_dataset <- get_sentiments("afinn")
sentiment_dataset <- arrange(sentiment_dataset, value)
```

```{r, message=FALSE, warning=FALSE}
sentiment <- trumpHashtags[,3:5] %>% unnest_tokens(output = 'word', input = 'text')

sentiment_dataset <- get_sentiments("afinn")
sentiment_dataset <- arrange(sentiment_dataset, -value)

#merge
sentiment <- merge(sentiment, sentiment_dataset, by = 'word')

#clean
sentiment$word <- NULL
sentiment$screen_name <- NULL

#get nearest hour of time for plot
sentiment$hour <- format(round(sentiment$created_at, units="hours"), format="%H:%M")

pivot <- sentiment %>%
group_by(hour) %>%
summarise(sentiment = mean(value))

#plot
ggplot(pivot[-1,], aes(x = hour, y = sentiment)) + geom_line(group = 1, col = "red") + geom_point() + geom_hline(yintercept=0) + theme_minimal() + labs(title = paste0('Average sentiment of tweetings mentioning "',search_term,'"'),
subtitle = paste0(pivot$hour[2],' - ',pivot$hour[nrow(pivot)],' on ', format(sentiment$created_at[1], '%d %B %Y')),
x = 'Date', y = 'Sentiment', caption = 'Source: Twitter API 10k Tweets')
ggsave("./Pictures/trumpTwitterSentiment.png")
```

```{r, message=FALSE, warning=FALSE}
bidenHashtags <-  readr::read_csv("./Data/joeBiden_TwitterAPI.csv")
search_term2 <- "#JoeBiden"
bidenHashtags$date <- substr(bidenHashtags$created_at,1,10)
bidenHashtags <- bidenHashtags[bidenHashtags$date == '2020-11-16',]
head(bidenHashtags)
```

```{r, message=FALSE, warning=FALSE}
by <- 'hour'
rtweet::ts_plot(bidenHashtags, by = by, trim = 1,col = "blue") + geom_point() + 
theme_minimal() + labs(title = paste0('Tweets mentioning "',
search_term2,'" by ',by),
x = 'Date', y = 'Count', caption = 'Source: Twitter API 10k Tweets')
ggsave("./Pictures/bidenByHour.png")
```

```{r, message=FALSE, warning=FALSE}
sentiment2 <- bidenHashtags[,3:5] %>% unnest_tokens(output = 'word', input = 'text')

#merge
sentiment2 <- merge(sentiment2, sentiment_dataset, by = 'word')

#clean
sentiment2$word <- NULL
sentiment2$screen_name <- NULL

#get nearest hour of time for plot
sentiment2$hour <- format(round(sentiment2$created_at, units="hours"), format="%H:%M")

pivot2 <- sentiment2 %>%
group_by(hour) %>%
summarise(sentiment2 = mean(value))

#plot
ggplot(pivot2[-1,], aes(x = hour, y = sentiment2)) + geom_line(group = 1, col = "blue") + geom_point() + geom_hline(yintercept=0) + theme_minimal() + labs(title = paste0('Average sentiment of tweetings mentioning "',search_term2,'"'),
subtitle = paste0(pivot2$hour[2],' - ',pivot2$hour[nrow(pivot2)],' on ', format(sentiment2$created_at[1], '%d %B %Y')),
x = 'Date', y = 'Sentiment', caption = 'Source: Twitter API 10k Tweets')
ggsave("./Pictures/bidenTwitterSentiment.png")
```

### 2.From fivethirtyeight prediction models of the election for 2020. Who is likely to win? "popular vote"

     After cleaning and filtering the fivethirtyeight dataset. We found the Biden is predicted to win in 2020 and also the data from 2020-2018. We filtered the fte_grade to be a C or better because we need reliable data to predicted who wold win for the popular vote
```{r, message=FALSE, warning=FALSE}
pollingData <- readr::read_csv("./Data/president_polls_fivethirtyeight.csv")
pollingData <- pollingData %>% filter(fte_grade != "NA")
pollingData <- pollingData %>% filter(fte_grade != "C-")
pollingData <- pollingData %>% filter(fte_grade != "D")
pollingData <- pollingData %>% filter(fte_grade != "D-")
pollingData <- pollingData %>% filter(fte_grade != "F") %>% select(poll_id, state,fte_grade, created_at, candidate_name, pct, sample_size) %>%
  filter(candidate_name == "Joseph R. Biden Jr." | candidate_name == "Donald Trump")

pollingData$year <- year(as.Date(strptime(pollingData$created_at, "%d/%m/%Y %H:%M")))

joeBidenPolls<- pollingData %>%
                filter(candidate_name == "Joseph R. Biden Jr.") %>%
                  summarise(
                  averagePollPrediction = sprintf("%0.2f",mean(pct))
                  ) 

donaldTrumpPolls<- pollingData%>%
                  filter(candidate_name == "Donald Trump") %>%
                  summarise(
                  averagePollPrediction = sprintf("%0.2f",mean(pct))
                  ) 

candidate_name <- c("Biden", "Trump")
averagePollPrediction <- c(joeBidenPolls$averagePollPrediction, donaldTrumpPolls$averagePollPrediction)
pollingSamplePrediction <-data.frame(candidate_name, averagePollPrediction)

pollingSamplePrediction$averagePollPrediction <- as.numeric(pollingSamplePrediction$averagePollPrediction)

ggplot(pollingSamplePrediction, aes(x = candidate_name, y = averagePollPrediction, fill = candidate_name)) + geom_bar(stat="identity", position = "dodge") + labs(title = 'Poll Surveys of Candidates since 2018',
       caption = "Data source: fivethirtyeight 2018-2020") + xlab('Candidate') + ylab('Average Polling Precinct') + scale_fill_manual("legend", values = c("Biden" = "#2641c9", "Trump" = "#c92626")) + geom_text(aes(label=averagePollPrediction), position=position_dodge(width=0.9), vjust=-0.25, size = 6) + ylim(0, 100)
ggsave("./Pictures/predicted2018-2020.png")

joeBidenPolls2020<- pollingData %>%
                filter(candidate_name == "Joseph R. Biden Jr." & year == "20") %>%
                  summarise(
                  averagePollPrediction = sprintf("%0.2f",mean(pct))
                  ) 

donaldTrumpPolls2020<- pollingData%>%
                  filter(candidate_name == "Donald Trump" & year == "20") %>%
                  summarise(
                  averagePollPrediction = sprintf("%0.2f",mean(pct))
                  ) 

candidate_name <- c("Biden", "Trump")
averagePollPrediction2020 <- c(joeBidenPolls2020$averagePollPrediction, donaldTrumpPolls2020$averagePollPrediction)
pollingSamplePrediction2020 <-data.frame(candidate_name, averagePollPrediction2020)

pollingSamplePrediction$averagePollPrediction <- as.numeric(pollingSamplePrediction2020$averagePollPrediction)

ggplot(pollingSamplePrediction, aes(x = candidate_name, y = averagePollPrediction, fill = candidate_name)) + geom_bar(stat="identity", position = "dodge") + labs(title = 'Poll Surveys of Candidates of 2020',
       caption = "Data source: fivethirtyeight 2020") + xlab('Candidate') + ylab('Average Polling Precinct') + scale_fill_manual("legend", values = c("Biden" = "#2641c9", "Trump" = "#c92626")) + geom_text(aes(label=averagePollPrediction), position=position_dodge(width=0.9), vjust=-0.25, size = 6) + ylim(0, 100)
ggsave("./Pictures/predicted2020.png")

```

### 3. From kcore-analytics prediction models of the election for 2020. Who is likely to win?
```{r, message=FALSE, warning=FALSE}
bidenElectoralVotes <- poll %>%
                            filter(bidenPercent > trumpPercent)
trumpElectoralVotes <- poll %>%
                            filter(trumpPercent > bidenPercent)
name <- c('Trump', 'Biden')
totalVotes <- c(sum(trumpElectoralVotes$electoralVotes), sum(bidenElectoralVotes$electoralVotes))

candidatePredictions <- data.frame(name, totalVotes)

ggplot(candidatePredictions, aes(x = name, y = totalVotes, fill = name)) + geom_bar(stat="identity", position = "dodge") + ggtitle('Total Number of Predicted Electoral Votes per Candidate') + xlab('Candidate') + ylab('Total Number of Predicted Electoral Votes') + scale_fill_manual("legend", values = c("Biden" = "#2641c9", "Trump" = "#c92626")) + geom_text(aes(label=totalVotes), position=position_dodge(width=0.9), vjust=-0.25, size = 6) + ylim(0, 300)
```

    By this graph, we can predict that Biden is likely to win 277-261 votes. Although the actual counts differ than what it turned out to be, the prediction for the president turned out to be correct.
  
```{r, message=FALSE, warning=FALSE}
poll$winner <- ifelse(poll$bidenPercent > poll$trumpPercent, 1, 0)
poll$region <- tolower(poll$state)
states <- map_data('state')
state_win <- data.frame(poll$region, poll$winner)
state_win2 <- full_join(states, state_win, by = c('region' = 'poll.region'))

ggplot(state_win2, aes(x = long, y = lat)) + 
  geom_polygon(aes(group = group)) +
  geom_polygon(state_win2, mapping = aes(x = long, y = lat, group = group, fill = ifelse(poll.winner == 1, "#2641c9", "#c92626"))) + 
  geom_path(aes(group = group), color = 'white') + 
  scale_fill_manual(labels = c('Biden', 'Trump', ''), values=c("#2641c9", "#c92626")) +
  coord_map() + 
  ggtitle('Predicted Voted Political Party by State') + xlab('') + ylab('') + labs(fill = 'Predicted State Winner') + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```
    
    Based off of this map, we can tell that the states each candidate is predicted to win is pretty similar to what they normally are. However, Biden is predicted to flip some states out west, in the midwest, and a couple of states in the east.
    
### 4. What States are Currently Republican or Democrat?

```{r, message=FALSE, warning=FALSE}
# Creating variable lists
region <- c("alabama","alaska", "arizona", "arkansas", "california",  "colorado", "connecticut", "delaware",
            "florida", "georgia","hawaii","idaho","illinois","indiana","iowa","kansas","kentucky","louisiana",
           "maine","maryland", "massachusetts", "michigan", "minnesota", "mississippi",
           "missouri","montana","nebraska","nevada","new hampshire","new jersey","new mexico",
           "new york","north carolina","north dakota","ohio","oklahoma","oregon","pennsylvania","rhode island",
           "south carolina","south dakota","tennessee","texas","utah","vermont","virginia",
           "washington","west virginia","wisconsin","wyoming")

vote <- c(2,2,1,2,1,1,1,1,2,1,1,2,1,2,2,2,2,2,0,1,1,
          1,1,2,2,2,3,1,1,1,1,1,2,2,2,2,1,1,1,2,2,2,2,2,1,1,1,2,1,2)
```

    3 - Nebraska split
    2 - Trump
    1 - Biden
    0 - Maine split

```{r, message=FALSE, warning=FALSE}
# Making the map
state_decision <- data.frame(region, vote)
state_votes <- full_join(states, state_decision)
```

    Shown above are the states each candidate won during this election. Looking back at the predicted map, you can see that the predicted map was pretty similar to what the outcome actually ended up being.

```{r, message=FALSE, warning=FALSE}
ggplot(state_votes, aes(x = long, y = lat)) + 
  geom_polygon(aes(group = group)) +
  geom_polygon(state_votes, mapping = aes(x = long, y = lat, group = group, fill = factor(vote))) + 
  geom_path(aes(group = group), color = 'white') + 
  scale_fill_manual(labels = c('Biden/Trump Split', 'Biden', 'Trump', 'Trump/Biden Split', ''), values=c('#617bfa', "#2641c9", "#c92626", '#f07878')) +
  coord_map() + 
  ggtitle('Voted Political Party by State') + xlab('') + ylab('') + labs(fill = 'Total Accidents') + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
```

#### Importing Datasets and Loading Stop Words

```{r, message=FALSE, warning=FALSE}
biden <- read.csv("https://raw.githubusercontent.com/elbalke/ds202_finalproject/main/Data/JoeBidenTweets.csv", stringsAsFactors = FALSE)

trump <- read.csv("https://raw.githubusercontent.com/elbalke/ds202_finalproject/main/Data/realdonaldtrump.csv", stringsAsFactors = FALSE)

# Loading stop words (irrelevant words such as "is" "are" "the" etc.)
data("stop_words")
```

### 5. What topics do both canidates talk about?

#### Biden Topics

```{r, message=FALSE, warning=FALSE}
# Removing punctuation from tweets
biden$cleanTxt <- gsub("http.*","",   biden$tweet)
biden$cleanTxt <- gsub("https.*","",   biden$tweet)
bidenTweets <- biden$cleanTxt
```

```{r, message=FALSE, warning=FALSE}
# Breaking tweets into words and removing stop words
cleanedBidenTweets <- biden %>%
  dplyr::select(cleanTxt) %>%
  unnest_tokens(word, cleanTxt)

cleanedBidenWords <- cleanedBidenTweets %>%
  anti_join(stop_words)

# Removing non-words
cleanedBidenWords2 <- subset(cleanedBidenWords, word!="http" & word!="t.co")
```

```{r, message=FALSE, warning=FALSE}
# Visualizing most common words
cleanedBidenWords2 %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col(fill = "blue") +
  coord_flip() +
  ggtitle("Most Common Words in Joe Biden Tweets") + 
  xlab("Word") +
  ylab("Count")
```

#### Trump Topics

```{r, message=FALSE, warning=FALSE}
# Removing punctuation from tweets
trump$cleanTxt <- gsub("http.*","",   trump$content)
trump$cleanTxt <- gsub("https.*","",   trump$content)
trumpTweets <- trump$cleanTxt
```

```{r, message=FALSE, warning=FALSE}
# Breaking tweets into words and removing stop words
cleanedTrumpTweets <- trump %>%
  dplyr::select(cleanTxt) %>%
  unnest_tokens(word, cleanTxt)

cleanedTrumpWords <- cleanedTrumpTweets %>%
  anti_join(stop_words)

# Removing non-words
cleanedTrumpWords2 <- subset(cleanedTrumpWords, word!="http" & word!="pic.twitter.com" & word!="bit.ly")
```

```{r, message=FALSE, warning=FALSE}
# Visualizing most common words
cleanedTrumpWords2 %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col(fill = "red") +
  coord_flip() +
  ggtitle("Most Common Words in Donald Trump Tweets") + 
  xlab("Word") +
  ylab("Count")
```

#### Common Topics

```{r, message=FALSE, warning=FALSE}
# Finding common topics using inner_join
commonWords <- inner_join(cleanedBidenWords2, cleanedTrumpWords2, by = "word")
```

```{r, message=FALSE, warning=FALSE}
# Visualizing common topics
commonWords %>%
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col(fill = "purple") +
  ggtitle("Common Topics Between Biden and Trump") + 
  xlab("Topic") +
  ylab("Count")
```

    From this bar chart we can identify two major topics that both Biden and Trump are tweeting about.  The most talked about topic is Donald Trump.  Trump likes to talk about himself and Biden often criticizes him, so it's no surprise the words "trump" and "donald" rank 1st and 3rd respectively.  The next major topic both people talk about is the president.  The data for Biden's tweets begins in 2007 and Trump's begins in 2009, so the topic of president has changed over time.

## 6.Who has more retweets and favorites for their tweets?
```{r}
bidenTweets <- read.csv("./Data/JoeBidenTweets.csv")
trumpTweets <- read.csv("./Data/realdonaldtrump.csv")
bidenGroup <- bidenTweets %>% summarise(
                          BidenTotalRetweets = sum(bidenTweets$retweets),
                          BidenTotalFavorites = sum(likes),
                          BidenAverageRetweets = mean(retweets),
                          BidenAverageFavorites = mean(likes),
                          BidenNumTweets = n())
trumpGroup <- trumpTweets %>% summarise(
                          TrumpTotalRetweets = sum(retweets),
                          TrumpTotalFavorites = sum(favorites),
                          TrumpAverageRetweets = mean(retweets),
                          TrumpAverageFavorites = mean(favorites),
                          TrumpNumTweets = n()
)

totalRetweets <- c(trumpGroup$TrumpTotalRetweets, bidenGroup$BidenTotalRetweets)
totalFavorites <- c(trumpGroup$TrumpTotalFavorites, bidenGroup$BidenTotalFavorites)
averageRetweets <- c(trumpGroup$TrumpAverageRetweets, bidenGroup$BidenAverageRetweets)
averageFavorites <- c(trumpGroup$TrumpAverageFavorites, bidenGroup$BidenAverageFavorites)
numberOfTweets <- c(trumpGroup$TrumpNumTweets, bidenGroup$BidenNumTweets)
candidatesTweetData <- data.frame(name, totalRetweets, totalFavorites, averageRetweets, averageFavorites, numberOfTweets)

ggplot(candidatesTweetData, aes(x = name, y = numberOfTweets, fill = name)) + geom_bar(stat="identity", position = "dodge") + ggtitle('Total Number of Tweets per Candidate') + xlab('Candidate') + ylab('Total Number of Tweets') +
scale_fill_manual("legend", values = c("Biden" = "#2641c9", "Trump" = "#c92626")) + geom_text(aes(label=numberOfTweets), position=position_dodge(width=0.9), vjust=-0.25, size = 6) + ylim(0, 45000)

ggplot(candidatesTweetData, aes(x = name, y = totalRetweets, fill = name)) + geom_bar(stat="identity", position = "dodge") + ggtitle('Total Number of Retweets per Candidate') + xlab('Candidate') + ylab('Total Number of Retweets') + scale_fill_manual("legend", values = c("Biden" = "#2641c9", "Trump" = "#c92626")) + geom_text(aes(label=totalRetweets), position=position_dodge(width=0.9), vjust=-0.25, size = 6) + ylim(0, 280000000)

ggplot(candidatesTweetData, aes(x = name, y = totalFavorites, fill = name)) + geom_bar(stat="identity", position = "dodge") + ggtitle('Total Number of Favorites per Candidate') + xlab('Candidate') + ylab('Total Number of Favorites') + scale_fill_manual("legend", values = c("Biden" = "#2641c9", "Trump" = "#c92626")) + geom_text(aes(label=totalFavorites), position=position_dodge(width=0.9), vjust=-0.25, size = 6) + ylim(0, 1175000000)

ggplot(candidatesTweetData, aes(x = name, y = averageRetweets, fill = name)) + geom_bar(stat="identity", position = "dodge") + ggtitle('Average Number of Retweets per Candidate') + xlab('Candidate') + ylab('Average Number of Retweets') + scale_fill_manual("legend", values = c("Biden" = "#2641c9", "Trump" = "#c92626")) + geom_text(aes(label=round(averageRetweets, digits = 2)), position=position_dodge(width=0.9), vjust=-0.25, size = 6) + ylim(0, 6400)

ggplot(candidatesTweetData, aes(x = name, y = averageFavorites, fill = name)) + geom_bar(stat="identity", position = "dodge") + ggtitle('Average Number of Favorites per Candidate') + xlab('Candidate') + ylab('Average Number of Favorites') + scale_fill_manual("legend", values = c("Biden" = "#2641c9", "Trump" = "#c92626")) + geom_text(aes(label=round(averageFavorites, digits = 2)), position=position_dodge(width=0.9), vjust=-0.25, size = 6) + ylim(0, 34000)
```

    Based on this data and the graphs, Trump has a significant more amount of retweets and favorites than Biden. However, Trump also has 37,576 more tweets than Biden. We looked into when the data sets were most recently updated and what date range the tweets were collected. The dataset we used for Trump's tweets had tweets from May 4th, 2009 through June 30th, 2020. There are 9 total years in this dataset. The dataset we used for Biden's tweets had tweets from April 9th, 2012 through April 30th, 2020 along with two tweets from 2007. There are 9 total years in this dataset. Although Trump's dataset started earlier, most of the tweets were from 2011 on and since Trump tweets an average of 3,612.67 times a year and Biden tweets an average of 641.78 times a year, I don't think having another year of Biden's tweets would make a big enough of a difference to get the counts closer to each other. Below will be an analyses of the average tweets per year each candidate tweeted. Although Trump has tweeted many more times, he barely has more average retweets per tweet than Biden and Biden actually has more average favorites than Trump.

## 7.What are the common hashtags used between Biden and Trump?

```{r}
trumpHashtags <- unlist(strsplit(trumpTweets$hashtags, ','))
trumpHashtags <- tibble(trumpHashtags)
allTrumpHashtags <- trumpHashtags
trumpHashtags <- unique(trumpHashtags$trumpHashtags)
trumpHashtags <- tibble(trumpHashtags)

bidenHashtags <- unlist(strsplit(bidenTweets$hashtags, ','))
bidenHashtags <- tibble(bidenHashtags)
allBidenHashtags <- bidenHashtags
bidenHashtags <- unique(bidenHashtags$bidenHashtags)
bidenHashtags <- tibble(bidenHashtags)
hashtagJoin <- inner_join(trumpHashtags, bidenHashtags, by = c('trumpHashtags' = 'bidenHashtags'))
colnames(hashtagJoin) <- c('hashtag')
print(as_tibble(hashtagJoin), n = 29)

colnames(allBidenHashtags) <- c('hashtag')
colnames(allTrumpHashtags) <- c('hashtag')
allHashtags <- rbind(allBidenHashtags, allTrumpHashtags)
allHashtags <- allHashtags %>%
                    group_by(hashtag) %>%
                    summarise(timesUsed = n())
top5Hashtags <- inner_join(allHashtags, hashtagJoin, by = c('hashtag')) %>%
                        arrange(-timesUsed) %>%
                        head(5)
ggplot(top5Hashtags, aes(x = hashtag, y = timesUsed, fill = hashtag)) + geom_bar(stat="identity") + coord_flip() +  ggtitle('Times Used Between Candidates') + xlab('hashtag') + ylab('Top Five Hashtags Used Between the Two Candidates')
```

    Based off of these findings, it seems like the majority of commonalities between the two candidates' hashtag usage is that most of them seem to either be a holiday, a day celebrating a person or group of people, standing in solidarity after a tragedy (for example #VegasStrong), or an election. Although #DemDebate is the most popular hashtag, with over 220 usages between the two candidates, Biden has used that hashtag 201 times so he has used it a majority of the time.
   