---
title: "calebpurcell.Rmd"
author: "Caleb Purcell"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Packages

```{r, message = FALSE, warning = FALSE}
library(ggplot2)
library(tidyverse)
library(tidytext)
library(textdata)
library(dplyr)
library(plotly)
library(lubridate)
```

## Importing Datasets and Loading Stop Words

```{r}
biden <- read.csv("https://raw.githubusercontent.com/elbalke/ds202_finalproject/main/JoeBidenTweets.csv", stringsAsFactors = FALSE)

trump <- read.csv("https://raw.githubusercontent.com/elbalke/ds202_finalproject/main/realdonaldtrump.csv", stringsAsFactors = FALSE)

# Loading stop words (irrelevant words such as "is" "are" "the" etc.)
data("stop_words")
```

## 5. What topics do both canidates talk about?

### Biden Topics

```{r}
# Removing punctuation from tweets
biden$cleanTxt <- gsub("http.*","",   biden$tweet)
biden$cleanTxt <- gsub("https.*","",   biden$tweet)
bidenTweets <- biden$cleanTxt
```

```{r, message = FALSE}
# Breaking tweets into words and removing stop words
cleanedBidenTweets <- biden %>%
  dplyr::select(cleanTxt) %>%
  unnest_tokens(word, cleanTxt)

cleanedBidenWords <- cleanedBidenTweets %>%
  anti_join(stop_words)

# Removing non-words
cleanedBidenWords2 <- subset(cleanedBidenWords, word!="http" & word!="t.co")
```

```{r, message = FALSE}
# Visualizing most common words
cleanedBidenWords2 %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col(fill = "blue") +
  coord_flip() +
  ggtitle("Most Common Words in Joe Biden Tweets") + 
  xlab("Word") +
  ylab("Count")
```

### Trump Topics

```{r}
# Removing punctuation from tweets
trump$cleanTxt <- gsub("http.*","",   trump$content)
trump$cleanTxt <- gsub("https.*","",   trump$content)
trumpTweets <- trump$cleanTxt
```

```{r, message = FALSE}
# Breaking tweets into words and removing stop words
cleanedTrumpTweets <- trump %>%
  dplyr::select(cleanTxt) %>%
  unnest_tokens(word, cleanTxt)

cleanedTrumpWords <- cleanedTrumpTweets %>%
  anti_join(stop_words)

# Removing non-words
cleanedTrumpWords2 <- subset(cleanedTrumpWords, word!="http" & word!="pic.twitter.com" & word!="bit.ly")
```

```{r, message = FALSE}
# Visualizing most common words
cleanedTrumpWords2 %>%
  count(word, sort = TRUE) %>%
  top_n(20) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col(fill = "red") +
  coord_flip() +
  ggtitle("Most Common Words in Donald Trump Tweets") + 
  xlab("Word") +
  ylab("Count")
```

### Common Topics

```{r}
# Finding common topics using inner_join
commonWords <- inner_join(cleanedBidenWords2, cleanedTrumpWords2, by = "word")
```

```{r, message = FALSE}
# Visualizing common topics
commonWords %>%
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = word, y = n)) +
  geom_col(fill = "purple") +
  ggtitle("Common Topics Between Biden and Trump") + 
  xlab("Topic") +
  ylab("Count")
```

    From this bar chart we can identify two major topics that both Biden and Trump are tweeting about.  The most talked about topic is Donald Trump.  Trump likes to talk about himself and Biden often criticizes him, so it's no surprise the words "trump" and "donald" rank 1st and 3rd respectively.  The next major topic both people talk about is the president.  The data for Biden's tweets begins in 2007 and Trump's begins in 2009, so the topic of president has changed over time.